---
title: "IMDB TOP 250 Movies"
author: "Hamed Ahmadi"
institute: "American University"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, rladies, rutgers-fonts]
    nature:
      highlightStyle: github
      highlightLines: true
      ratio: "16:9"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  prompt = FALSE,
  tidy = FALSE, 
  comment = NA,
  message = FALSE, 
  warning = FALSE
)
library(tidyverse)
library(patchwork)
df <- read_csv("IMDB_Top250Engmovies2_OMDB_Detailed.csv")
```
class: left, middle
## Data

The dataset encompasses the top 250 movies from IMDB and features detailed attributes, including genre, ratings, release year, runtime, director, studio, language, country, and actors, among others. The data, owned by "TheMitchWorksPro" and sourced from https://data.world/studentoflife/imdb-top-250-lists-and-5000-or-so-data-records, was curated to focus on genre trends and the distribution of ratings.
  
---
class: left
## IMDB Top 250 Ratings Distribution

```{r, fig.width=16, fig.height=8}
plot1 = ggplot(df, aes(x = imdbRating)) +
  geom_histogram(binwidth = 0.1, fill = '#9966CC', color = 'black') +
  theme_minimal() +
  labs(
    x = "IMDB Rating",
    y = "Count") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title = element_text(size =14),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
        panel.grid.minor = element_blank()  # Optional: Removes minor grid lines
)

plot1
```

---
## Top 250 Movies Across Decades

```{r combined_plot, message=FALSE, fig.width=14, fig.height=7}

# Assuming df is your data frame and it has been loaded correctly
df <- df %>%
  mutate(
    Year = as.numeric(Year),  # Ensure Year is numeric
    Decade = floor(Year / 10) * 10  # Create Decade column
  ) %>%
  separate_rows(Genre, sep = ", ")  # Separate genres into individual rows for each movie

# Get the ten most common genres
top_genres <- df %>%
  count(Genre) %>%
  top_n(10) %>%
  pull(Genre)

# Filter the data to include only the ten most common genres
df <- df %>% filter(Genre %in% top_genres)

# Define a custom color palette with distinct colors for each genre
genre_colors <- c("#E41A1C", "#377EB8", "#4DAF4A", "#FF7F00", "#F781BF", 
                  "#A65628", "#984EA3", "#999999", "#66C2A5", "#FFD92F")

# Now, calculate the proportion of movies within each genre for each decade
df_decade_genre <- df %>%
  group_by(Decade, Genre) %>%
  summarise(MovieCount = n(), .groups = 'drop') %>%
  group_by(Decade) %>%
  mutate(TotalMovies = sum(MovieCount),
         Proportion = MovieCount / TotalMovies) %>%
  ungroup()


plot2_1 = ggplot(df_decade_genre, aes(x = as.factor(Decade), y = Proportion, fill = Genre)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_fill_manual(values = genre_colors) +
  theme_minimal() +
  labs(title = "Genre Prevalence in Top 250 Movies Across Decades",
       x = "Decade",
       y = "Proportion of Movies",
       fill = "Genre") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = function(x) paste0(x, "s")) + # Adding "s" to decade labels
  theme(
    plot.title = element_text(hjust = 0, size = 20, face = "bold"), # Center and enhance title size for better readability
    legend.position = "right", # Adjust legend to the right side
    legend.title = element_text(size = 16), # Enhance legend title size
    legend.text = element_text(size = 14), # Enhance legend text size
    legend.key.size = unit(0.9, "cm"),
    axis.title.x = element_blank(), # Enhance X axis title size
    axis.title.y = element_text(size = 12), # Enhance Y axis title size
    axis.text.x = element_text(angle = 45, hjust = 1, size = 14, face = "bold"), # Rotate X axis labels for better legibility
    axis.text.y = element_text(size = 12) # Enhance Y axis text size
  )



df_decade_count <- df %>%
  group_by(Decade) %>%
  summarise(MovieCount = n(), .groups = 'drop') %>%
  arrange(Decade) # Ensure decades are in chronological order

# Enhanced visualization:
plot2_2 = ggplot(df_decade_count, aes(x = as.factor(Decade), y = MovieCount)) +
  geom_bar(stat = "identity", fill = 'orange', color = 'darkgrey') + # Improved color for better visual appeal
  geom_text(aes(label = MovieCount), vjust = -0.3, color = "black", size = 6) + # Add data labels for clarity
  theme_minimal(base_size = 12) + # Increase base font size for better readability
  labs(title = "Top 250 Movies Over Decades",
       x = "Decade",
       y = "Number of Movies") +
  scale_x_discrete(labels = function(x) paste0(x, "s")) + # Enhance x-axis labels for clarity
  theme(plot.title = element_text(hjust = 0, size = 20, face = "bold"), # Center and increase the title font size
        plot.subtitle = element_text(hjust = 0, size = 12), # Center and adjust subtitle font size
        axis.title.x = element_blank(), # Enhance X axis title size
    axis.title.y = element_text(size = 12), # Enhance Y axis title size
    axis.text.x = element_text(angle = 45, hjust = 1, size = 14, face = "bold"), 
    axis.text.y = element_text(size = 12), 
    panel.grid.minor = element_blank(),  # Optional: Removes minor grid lines
        legend.position = "none")



# Combine the plots
combined_plot = plot2_2 + plot2_1

# Display the combined plot
combined_plot

```

